### 2023-07-18 submission_4/submission_4.qmd

---
title: "submission_4"
subtitle: "Documentation"
subsubtitle: "Leaf traits measurements"
author: | 
    | **Group 4** 
    | Denis Streitmatter 
    | Marie-Louise Korte 
    | Louis Trinkle 
    | Maxwell Sivertsen 
    | Fabian Götz 
    | Lukas Erzfeld
date: last-modified
format:
    pdf:
        documentclass: scrreport
        theme: united
        toc: true
        toc-depth: 2
        number-sections: false 
        number-depth: 1
        toc-title: Table of Contents
        toc-location: body
        colorlinks: true
header-includes:  
- \usepackage{graphicx}
- \usepackage{lscape}
- \usepackage{array}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
{{< pagebreak >}}

#