### 2023-07-19 submission_4/submission_4.qmd
Adds sources for subission 4

---
title: "submission_4"
subtitle: "Documentation"
subsubtitle: "Leaf traits measurements"
author: | 
    | **Group 4** 
    | Denis Streitmatter 
    | Marie-Louise Korte 
    | Louis Trinkle 
    | Maxwell Sivertsen 
    | Fabian Götz 
    | Lukas Erzfeld
date: last-modified
format:
    pdf:
        documentclass: scrreport
        theme: united
        toc: true
        toc-depth: 2
        number-sections: false 
        number-depth: 1
        toc-title: Table of Contents
        toc-location: body
        colorlinks: true
header-includes:  
- \usepackage{graphicx}
- \usepackage{lscape}
- \usepackage{array}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
{{< pagebreak >}}

# Submission 4: TLS and Sentinel 2 comparison

## General study concept and conditions

The purpose of this document is the comparison of terrestial laser/LiDAR scanning and Sentinel 2 (satellite) data of the same spot.

**Insert some description and the data of the plot. I wasn't there, so probably someone else should do it**

## Algorithm idea

## Results

This section lists, describes and tries to explain the plots and graphs generated by the algorithm described in the previous section.

### Tree height statistic comparison

In this section the general measurements are compared.

*First the general heigths:*

![Height calculation of TLS (resampled to resolution of the sentinel data)](TLS_plots/figures/chm_height.png)

### Residual computation

In this section we compute the residuals of the